\bf Drift Removal for Time Series Data Using Quantile Trend Filtering

### 2019-01-28 Manuscript/detrendify_JCGS.tex
Updated simulation study text

	\begin{figure}
		\label{fig:trendfilter}
		\centering
		\caption{Examples of trend filtering solutions. The true trend is shown in black, while the estimated trend is shown in red.}
		\includegraphics[width = 0.45\linewidth]{Figures/trend_filter_eg1.png}
		\includegraphics[width = 0.45\linewidth]{Figures/trend_filter_eg2.png}
	\end{figure}

@@

	\begin{figure}[b!]
		\label{fig:raw_spod}
		\caption{Example of 3 co-located SPod PID sensor readings.}
		\includegraphics[width = \linewidth]{Figures/uncorrected_data.png}
	\end{figure}

	In this application, as well as those described in \cite{Ning2014}, \cite{marandi2015qualitative}, and \cite{pettersson2013algorithm}, it is not the mean trend that is desired but rather the baseline trend, which we can think of as the trend in a low quantile of the data.  \cite{Koenker1978} was the first to propose substituting the squared loss used in regression with the check loss function  (Eq. \ref{eq1}) to estimate a conditional quantile instead of the conditional mean.

	\begin{equation}
	\label{eq1}
	 \rho_{\tau}(z) = \sum_{i=1}^n z_i(1-\mathbf{I}(z_i<0))
	\end{equation}

@@

	The number of parameters to be estimates is equal to the number of observations multiplied by the number of quantiles of interest. As the size of the data and the number of quantiles grows, all solvers will eventually break.

@@

	\subsection{Peak Detection}
	We use another simulation design based on the applied problem we aim to solve. We assume that the measured data can be represented by 
	\begin{equation}
	Y(t) = s(t) + b(t) + \epsilon(t)
	\end{equation} 
	where $s(t)$ is the true signal at time $t$, $b(t)$ is the drift component that varies smoothly over time and $\epsilon(t) \sim N(0, \sigma^2)$ is an error component. We assume $t$ is a uniformly spaced sequence of time points between 0 and 1. We generate $b(t)$ using a cubic natural spline basis function with degrees of freedom sampled from $n/50$ to $n/25$ with equal probability, and coefficients drawn from an exponential distribution with rate 1. The true signal function is assumed to be zero with Gaussian peaks. The number of peaks is sampled from $n/100$ to $n/50$ with equal probability with centers uniformly distributed between 0.1 and 0.9 and bandwidths uniformly distributed between $1/n$ and $5/n$ and areas uniformly distributed between 0 and $20/n$. One hundred datasets were generated for $n=\{300, 500, 1000, 5000\}$. We compare the methods ability to estimate the true quantiles of $b(t) + \epsilon$  for $\tau \in \{0.01, 0.05, 0.1\}$ and calculate the root mean squared error (RMSE).

	\begin{figure}[h]
		\caption{Example of simulated peaks, baseline, and observed measurements.}
		\includegraphics[width = \linewidth]{Figures/ex_peaks.png}
	\end{figure}

	In addition to the RMSE we also calculated the signal miss-classification rate. We first classified the ``true" peaks function into signal or not based on a threshold of 0.1. We then classified the detrended data using three different thresholds and calculated the fraction miss-classified. An illustration of the observations classified as signal after detrending compared to the ``true signal" is shown in Fig. 9.

@@

	\begin{figure}[h!]
		\caption{Example signal classification using threshold. Red indicates true signal $>0.1$, blue indicates classified as signal after baseline removal using eBIC detrendr and a threshold of 1.}
		\includegraphics[width = \linewidth]{Figures/peaks_eg_class.png}
	\end{figure}

>>>>

	\begin{figure}
		\centering
		\caption{Examples of trend filtering solutions. The true trend is shown in black, while the .estimated trend is shown in red.}
		\includegraphics[width = 0.45\linewidth]{Figures/trend_filter_eg1.png}
		\includegraphics[width = 0.45\linewidth]{Figures/trend_filter_eg2.png}
		\label{fig:trendfilter}
	\end{figure}

@@

	\begin{figure}[b!]
		\caption{Example of 3 co-located SPod PID sensor readings.}
		\includegraphics[width = \linewidth]{Figures/uncorrected_data.png}
		\label{fig:raw_spod}
	\end{figure}

	In this application, as well as those described in \cite{Ning2014}, \cite{marandi2015qualitative}, and \cite{pettersson2013algorithm}, it is not the mean trend that is desired but rather the baseline trend, which we can think of as the trend in a low quantile of the data.  \cite{Koenker1978} were the first to propose substituting the squared loss used in regression with the check loss function  (Eq. \ref{eq1}) to estimate a conditional quantile instead of the conditional mean.

	\begin{equation}
	\label{eq1}
	 \rho_{\tau}(z) = \sum_{i=1}^n z_i(1-\mathbf{I}(z_i<0)).
	\end{equation}

@@

	The number of parameters to be estimated is equal to the number of observations multiplied by the number of quantiles of interest. As the size of the data and the number of quantiles grows, all solvers will eventually break.

@@

	\subsection{Peak Detection}
	\begin{figure}[h]
		\caption{Example of simulated peaks, baseline, and observed measurements.}
		\includegraphics[width = \linewidth]{Figures/ex_peaks.png}
	\end{figure}

	We use another simulation design based on the applied problem we aim to solve. We assume that the measured data can be represented by 
	\begin{equation}
	Y(t) = s(t) + b(t) + \epsilon(t)
	\end{equation} 
	with $t = 1, ..., n$, where $s(t)$ is the true signal at time $t$, $b(t)$ is the drift component that varies smoothly over time and $\epsilon(t) \sim N(0, \sigma^2)$ is an error component. We generate $b(t)$ using a cubic natural spline basis function with degrees of freedom sampled from a Poisson distribution with mean parameter equal to $n/100$,  and coefficients drawn from an exponential distribution with rate 1. The true signal function is assumed to be zero with peaks generated using the Gaussian density function. The number of peaks is sampled from a binomial distribution with size equal to $n$ and probability equal to $0.005$ with location parameters uniformly distributed between $1$ and $n-1$ and bandwidths uniformly distributed between $2$ and $12$. The peaks were then multiplied by a factor that was randomly drawn from a normal distribution with mean 20 and standard deviation of 4. One hundred datasets were generated for $n=\{500, 1000, 2000, 4000\}$. We compare the ability of the methods to estimate the true quantiles of $b(t) + \epsilon$  for $\tau \in \{0.01, 0.05, 0.1\}$ and calculate the RMSE (Fig. \ref{fig:peaks_rmse}). 	
	\begin{figure}
		\caption{RMSE by method, quantile and data size for peaks design.}
		\includegraphics[width = \linewidth]{Figures/peaks_mse.png}	
		\label{fig:peaks_rmse}
	\end{figure}

	In our application, we want to accurately classify the observations into signal and no signal using a threshold. To evaluate the accuracy of our method compared to other methods we define true signal as any time when the simulated peak value is greater than 0.5. We compared three different trends for the baseline estimation and four different thresholds for separating the estimated signal after subtracting the estimated baseline from the observations.  An illustration of the observations classified as signal after detrending compared to the ``true signal" is shown in Fig. \ref{fig:peaks_class_eg}.  To compare the resulting signal classifications we calculate the class averaged accuracy (CAA). Defining $s(t) \in \{0,1\}$ as the true vector of signal classification and $\hat{s}(t) \in \{0,1\}$ as the estimated signal classification, the CAA is defined as
	\begin{equation}
	\mbox{CAA} = \frac{1}{2}\left(\frac{\sum_{t=1}^n \mathbf{I}[s(t) = 1 \cap \hat{s}(t)=1]}{\sum_{t=1}^n \mathbf{I}[s(t) = 1]} + \frac{\sum_{t=1}^n \mathbf{I}[s(t) = 0 \cap \hat{s}(t)=0]}{\sum_{t=1}^n \mathbf{I}[s(t) = 0]}\right).
	\end{equation}  
	We use this metric because our classes tend to be very un-balanced in this case with many more $0$s than $1$s. The CAA metric will always give a score of 0.5 for random guessing and also for trivial classifiers such as $\hat{s}(t) = 0$ for all $t$.

@@

	\begin{figure}[h!]
		\caption{Example signal classification using threshold. Red indicates true signal $>0.5$, blue indicates observations classified as signal after baseline removal using eBIC detrendr and a threshold of 1.2.}
		\includegraphics[width = \linewidth]{Figures/peaks_eg_class.png}
		\label{fig:peaks_class_eg}
	\end{figure}
